DRAFT: Harmonized and trait aggregation paper

### 2020-11-23 Paper/Draft_harmonization_paper.tex
23.11: Polishing/minor changes

In this study, we follow the terminology proposed by Schmera et al., where a grouping feature is defined as a general property (e.g. feeding mode) that comprises a "group of related traits (e.g., predator, shredder, etc.) that vary among species or among individuals within a species" \cite{schmera_proposed_2015}. Thus, we use the term grouping feature in place of which many studies use the term "trait", and the term trait instead of "trait state", "modality" or "trait category". Traits can be described using different codings (binary, fuzzy) that represent the prevalence of a characteristic in an organism.
% TODO: bring affinity/affinity scores in
% Could say that we call the proportionate description (0-1) "affinity"

@@

% Alternative from Ben:
Traits can be described in a binary fashion, or with multiple categories, ignoring uncertainty how the trait is expressed in any particular organism (e.g. adult terrestrial stage, presence of gills) or continuous (e.g. tolerance of pollution, body size). One approach for dealing with uncertainty is the use of fuzzy coded variables, where the traits are assigned probabilistic values. Fuzzy codes are used to account for plasticity in traits, variability in traits within taxonomic groups above species, and incomplete knowledge and are usually converted to proportions. In this study we denote the proportionate description (between 0-1) of traits as affinities or affinity scores.

However, invertebrate trait databases are heterogeneous regarding the coding they use for their traits \cite{culp_incorporating_2011}. Brown et al. \cite{brown_functional_2018} harmonized grouping features based on trait databases from Europe, North America, and New Zealand because in these trait databases identical grouping features are classified differently into traits. As the traits from North America were coded binary in contrast to the traits from Europe and New Zealand which have been established using affinity scores, the authors consulted experts to assign fuzzy coded traits to North American taxa or inferred them from the European trait database. Thus, it becomes apparent that using invertebrate trait data from several regions requires extensive data processing prior to the actual data analysis. A centralized database with standardized and unambiguous traits and a consistent coding of traits would minimize the data processing effort.
Discrepancies in the taxonomic resolutions (e.g. species, genus or family) when linking observational taxonomic data to trait database represents another challenge. When observations are on a more precise taxonomic level than data available in the trait databases (e.g. observations on species-level, trait data on genus-level) trait information of the less precise taxonomic level is often assigned (e.g. \cite{szocs_effects_2014, vos_taxonomic_2017}). Conversely, if trait information is only available on more precise taxonomic levels than the observed taxa, traits are aggregated to a less precise taxonomic level, e.g. \cite{poff_functional_2006, szocs_effects_2014, piliere_a._f._h._importance_2016, aspin_extreme_2019}. Studies have used different methods of trait aggregation, e.g. the mean \cite{magliozzi_functional_2019}, median \cite{szocs_effects_2014} or the mode \cite{piliere_a._f._h._importance_2016}. Up to now, studies on how and to which extent different trait aggregation methods influence trait-based analysis are missing. However, related knowledge would inform future studies regarding the choice of the aggregation method.

@@

% The necessity to aggregate and harmonize trait data due to the discrepancies in trait definitions motivated us to compare different trait aggregation methods and evaluate whether the harmonization of trait data ...
% As a basis for our analysis, we (1) established 4 novel invertebrate trait datasets consisting of  7 grouping features based on trait databases from Europe, North America, Australia, and New Zealand. To evaluate different trait aggregation methods we (2) compared trait affinities obtained through different trait aggregation methods to trait affinities assigned at family-level by experts for two of the established trait datasets. % Mention simulation
% (3) To evaluate how harmonizing grouping features and aggregating traits can modify results of trait-environment relationships we re-analysed data on the effect of anthropogenic salinisation on biological traits by Szöcs et al. \cite{szocs_effects_2014} using trait data from the established harmonized European trait datasets. By comparison with the original analysis, we investigated how harmonizing and aggregating trait data can modify the outcome of trait-environment relationships. Finally, we (4) present an overview of discrepancies in trait definition between the used invertebrate trait datasets and discuss challenges of trait data synthesis.

@@

We used fuzzy coded traits for establishing our harmonized datasets unless data quality prohibited. In the latter case we used binary traits, i.e. categorical and continuous traits were converted into binary traits. For example, in the  freshwaterecology.info database, the classification of the trait voltinism accounts for different faunistic regions. Hence, entries such as "arctic" or "boreal" of e.g. the trait univoltine were substituted with a value of 1. Implicitly, we assumed for binary traits that a value of 1 and 0 corresponds to the highest and no affinity for a particular trait. Fuzzy codes are reported with different ranges in the trait databases (e.g. freshwaterecology.info 0 to 10, Tachet database 0 to 3 or 0 to 5). We standardized these to a range between 0 and 1 and converted trait affinities to percentages. Thus, fuzzy coded and binary traits were in the same range.

@@

To examine the influence of the taxonomic hierarchy and the trait variability on the outcome of the different trait aggregation methods we created three hypothetical examples of different taxonomic hierarchies. 
1) A family with an equal number of genera and species (5 genera each with 5 species respectively), denoted as \textit{sim\_base}.
2) A family where one genus has a much larger number of species than the other 4 genera (1 genus with 13 species, 4 genera with 3 species respectively), denoted as \textit{sim\_extreme}. 
3) A family where all genera have a different number of species (8, 2, 7, 3, 5), denoted as \textit{sim\_variation}. In total every family consisted of 25 species. To each hypothetical taxonomic hierarchy a hypothetical grouping feature with 3 traits was assigned. The 25 affinities for each trait were simulated by sampling from a truncated normal distribution with a mean value of 0.5 and 5 levels of standard deviations (0.2, 0.4, 0.6, 0.8, and 1) respectively to simulate different levels of trait variability. The truncated normal distribution was bound to 0 and 1. Simulated trait affinities were converted to percentages, similar to the data processing of the trait databases. The sampling was repeated 100 times for each standard deviation. Hence, in total 100 datasets for each level of trait variability were simulated, resulting in 12500 simulated trait affinities ($25$ species per family $* 5$ levels of trait variability $* 100$ replicates). All trait aggregation methods were applied to each simulated dataset. The results were compared based on the produced range of aggregated trait affinities between levels of trait variability and hypothetical taxonomic hierarchies. We also compared for each simulated dataset the differences in trait affinities obtained by each aggregation method.

@@

Definitions of grouping features and traits varied in their level of detail in the original trait databases. The freshwaterecology.info database, the Tachet database and the North American database (Twardochleb et al.) provided more detailed descriptions of their trait information compared to the North American (Vieira et al.) and New Zealand database. An exception is the Australian trait database which is a collection of seven specific trait datasets \cite{kefford_integrated_2020}. Thus, grouping features occur multiple times with varying differentiation into traits. Depending on the dataset trait information is described with more or less detail.

@@

Various codings of traits are used throughout the databases (e.g. binary, fuzzy, continuous). The freshwaterecology.info and Australian use different codings in their databases. Tachet and the New Zealand database exclusively use fuzzy coding. Both North American trait databases contain categorical grouping features that can be converted into traits using a binary coding (Table \ref{tab:trait_databases_coding_differentiation}). Binary coding represents a simple approach in which a taxon either expresses a trait or not. Fuzzy coding characterizes the affinity of an organism to exert a certain trait. It is used to account for plasticity in traits, e.g. taking into account that traits can change over the development time of an organism. Usually, fuzzy coded affinities are converted into proportional values. Continuous coding is used for traits like body size.
% Last sentences are a bit redundant

% None of the seven grouping features compared here has the same differentiation of traits across all databases (Table \ref{tab:trait_databases})

@@

% 10 unique comparisons, 100 datasets, 5 levels of trait variability, 3 sim scenarios  
A comparison of the results of the aggregation methods for each simulated data set showed that in most simulated datasets the differences between the aggregated traits were small. Only 1.42 \% (213 out of 15.000) of all comparisons showed a difference equal or greater than an absolute trait affinity of 0.1. Most of these differences occurred for the \textit{sim\_extreme} scenario (83.5 \%). The majority of the differences equal or above 0.1 were found between the aggregation methods \textit{direct\_agg\textsubscript{mean}} and \textit{stepwise\_agg\textsubscript{median}}, \textit{direct\_agg\textsubscript{median}} and \textit{stepwise\_agg\textsubscript{median}}, and \textit{stepwise\_agg\textsubscript{median}} and \textit{weighted\_agg} (Figure S\ref{fig:sim_indv_runs}). 
% Differences were rather small, largest difference 0.265 for sim_extreme (stepwise_median vs weighted_agg)

@@

\begin{figure}[H]
    \label{fig:rda_species_scores}
    \centering
    \includegraphics[width=16.5cm, height=10cm]{Species_scores_rda.png}
    \caption{Species scores obtained by RDA from the original analysis \cite{szocs_effects_2014}, using harmonized grouping features, and using harmonized grouping features with traits aggregated to family-level. The dots represent the individual species scores for each analysed trait. The violin plot shows the density estimate of the species scores. Gray vertical lines indicate the median of the obtained species scores.}
\end{figure}

@@

% Summarize briefly results and synthesize (with Simulation)  
Our results show that aggregation of traits by the median compared to aggregating by the mean is often closer to the assigned traits at family level, i.e. it yields less differing cases. However, in those cases where differences in the trait aggregation methods that use the median occurred, these differences were greater compared to the aggregation methods that used the mean. The different weighting methods, i.e. aggregating stepwise, direct or weighted, did not seem to have a strong impact, because the distribution of absolute trait affinity differences to assigned traits obtained by the \textit{direct\_agg\textsubscript{mean}}, \textit{stepwise\_agg\textsubscript{mean}} and \textit{weighted\_agg} were relatively similar in both tested datasets. This also applies to the distributions of absolute trait affinity differences obtained by \textit{direct\_agg\textsubscript{median}} and \textit{stepwise\_agg\textsubscript{median}}. 
Instead, using the mean or the median changed the distribution of differences in trait affinities between aggregated and assigned traits. A possible reason why the use of different weightings for taxa appears to have little impact on trait aggregation is that a considerable portion of taxa used have a low numbers of genera or species entries. From the taxa that were compared from the North American trait dataset to the assigned traits at family-level, 13.8 \% were resolved on family-level and 61.7 \% were resolved on genus-level. 52.1 \% had 5 or less genera and 12.8 \% contained just one genus (Figure S\ref{fig:tax_hierarchy_NOA}). In the Australian dataset 21.4 \% of the compared taxa had trait information resolved on family-level, 39.5 \% resolved on genus level, and 39.1 \% resolved on species-level. 67.73 \% of the taxa contain 5 or less genera and 39.6 \% just contained one genus (Figure S\ref{fig:tax_hierarchy_AUS}). The simulation results indicated that the choice of the aggregation method for small numbers of genera per family (5) can be important when one genus has a much larger number of species compared to other genera within a particular family. % How often in the trait datasets?
In general however, inspecting the simulated datasets revealed that often aggregated traits obtained from the different aggregation methods only differ slightly.

% Differences between datasets 
Observed differences between aggregated and assigned traits were higher for the North American trait dataset. This can be explained by the binary coding of the traits in the North American dataset. As a results, traits have been assigned affinities of either 1 or 0. Thus naturally, even after normalization, traits of a particular grouping feature have a high difference of trait affinities to each other compared to when they were coded using fuzzy codes.

% Of course all the comparisons have no "generellen Gültigkeisanspruch" - what's the true aggregated trait affinity characterising a particular family? Expert assignments maybe also biased based on experience with a particular group of taxa within a family.  
Evaluation of the differences between aggregated and assigned traits is difficult because it remains unclear what the true value of a particular trait for a particular family is. It is indeed questionable whether the aggregation of trait information on species or genus-level to a point estimate on family-level can be an adequate description of a particular trait at all, especially for traits with a high variability. In fact, traits can vary strongly on more precise taxonomic levels than the family-level. For example, Monaghan and Soares found a high dispersion of trait affinities for genera within families in the Tachet database for traits of the grouping features body size, flow preference, and reproduction \cite{monaghan_improving_2013}. Furthermore, studies focusing on the lability of traits, i.e. how much traits are unconstrained by phylogeny, found that traits of ecological preferences (e.g. thermal preference), body size, resistance forms, and to a lesser extent feeding mode are labile, and thus possibly highly variable \cite{poff_functional_2006,wilkes_traitbased_2020}. Thus, it is likely that assigned traits and aggregated traits at family-level will especially differ in affinities for the aforementioned traits. Our analysis could confirm this partially. The highest possible difference of 1 between aggregated trait affinities and assigned trait affinities (in total in 6.9 \% of all cases) for the Australian dataset occurred most often for body size large and medium (42.5 \% and 39.2 \% of all cases with a difference of 1, respectively) across all comparisons. Both traits had also the highest mean and standard deviation in the absolute difference of trait affinities across all tested comparisons (size medium: mean 0.5, sd 0.29; size large: mean 0.59, sd 0.32). For the taxa compared from the North American dataset, the highest possible difference in trait affinities occurred most often for the trait body size medium (15.8 \% of all cases with the maximum difference of 1), followed by the traits respiration via plastron and spiracle and respiration via tegument (both 13 \% of all cases with a maximum difference of 1). Respiration via plastron and spiracle showed the highest mean and standard deviation in the absolute difference of trait affinities across all comparisons (mean 0.71, sd 0.33), followed by respiration via tegument (mean 0.5, sd 0.34) and body size medium (mean 0.49, sd 0.3). 
% Check for those cases that deviate a lot -> how have the aggregation and assignements performed when looking back at the original data?

% What to do with these traits? Also briefly talk about traits that are more consistent?

% Outlook: combine trait information for priors like in species distribution models/population growth models,
% see Kindvater et al and Brown & Roff 
% ? Constructing priors from fuzzy coded traits

% Further ideas: 
%   - Differences across grouping features
%   - Still some similarities between compared datasets -> Where does the classification come from?
%     Few sources used in NOA also used in AUS:
%     AUS: Merrit & Cummins 1978/1984 for feeding mode for a few genera and families 
%     NOA: Merritt, Richard W., Cummins, Berg 2008 -> but not clear if for feeding mode as well; according to trait description "Cummins, K. W. Trophic Relations of Aquatic Insects" was used for feeding mode
%   - Could not include all data due to data gaps -> much needed to be filled -> would maybe change the picture?

@@

The re-analysis showed that neither using harmonised grouping features nor aggregated traits changed drastically the results of the original study. The traits that did not respond to the salinity gradient (i.e. not on the extremes of the conductivity axis) in the re-analysis were those closest to the criterion that defined when a trait was associated with either high or low salinity (mahalanobis distance greater than the 97.5 \% quantile of the Chi-square distribution, see methods), namely feeding mode shredder, respiration gills, and life cycle duration traits. During harmonisation the trait shredder has been amalgamated based on three traits (Miner, xylophagus, and shredder). As a result, the trait affinities in the original dataset have a higher mean and standard deviation than in the aggregated and harmonized datasets (Table S\ref{tab:SI_resp_traits_summary_stats}).
% TODO: Check naming of tables! 
Thus, the signal in the original data for the trait shredder has presumably been weakened by the harmonization, which might be the reason that shredders showed a less strong reaction to the salinity gradient.

@@

\begin{itemize}
  \item \textbf{aggregated vs assigned traits}
  \begin{itemize}
    \item There is no "true" value to compare to, i.e. assigned traits at genus or family level can be biased as well (i.e. the person who assigned the trait was thinking about particular species within a genus not all). Comparison of species-level traits in fwe and traits on genus-level in tachet would be interesting in this context? 
    \item  Suggestions from Co-Authors: \\
     $\rightarrow$ should include where the classifications come from \\
     $\rightarrow$ what do we really know because someone has watched the animals and published their behaviour; what is extrapolated knowledge? \\
     $\rightarrow$ are the differences between continents because of evolution or only because scientists have done more research in one of the continents?
  \end{itemize}
  \item \textbf{Influence of aggregation methods on trait-environment relationships}
  \begin{itemize}
    \item We re-analysed only a small-scale study, what would be different when applying a large-scale study?
  \end{itemize}
\end{itemize}

@@

% 2) Influence of aggregation methods on trait-environment relationships: 
% Harmonization: Minor influence on RDA species scores
% RDA species scores 
%   -> Normalized eigenvectors
%   -> dependent on nr. of variables? Nr. of response variables not really influential 
% We used only a small-scale study -> What's different to large-scale studies?
% Further aggr. after harmonization: no difference to the harmonized dataset
% ?Logical because trait aggr. methods themselves showed only minor differences
% Complex Stat. techniques developed -> cannot use full potential

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Supporting information %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{table}{0}
\setcounter{figure}{0}

>>>>

In this study, we follow the terminology proposed by Schmera et al., where a grouping feature is defined as a general property (e.g. feeding mode) that comprises a "group of related traits (e.g., predator, shredder, etc.) that vary among species or among individuals within a species" \cite{schmera_proposed_2015}. Thus, we use the term grouping feature in place of which many studies use the term "trait", and the term trait instead of "trait state", "modality" or "trait category". Traits can be described using different codings (binary, fuzzy) that represent the prevalence of a characteristic in an organism.

@@

Traits can be described in a binary fashion, or with multiple categories, ignoring uncertainty how the trait is expressed in any particular organism (e.g. adult terrestrial stage, presence of gills) or continuous (e.g. tolerance of pollution, body size). One approach for dealing with uncertainty is the use of fuzzy coded grouping features, where the traits are assigned probabilistic values. Fuzzy codes are used to account for plasticity in traits, variability in traits within taxonomic groups above species, and incomplete knowledge and are usually converted to proportions. In this study we denote the proportionate description (0-1) of traits as affinities or affinity scores.

However, invertebrate trait databases are heterogeneous regarding the coding they use for their traits \cite{culp_incorporating_2011}. For example, in the study from Brown et al. \cite{brown_functional_2018} the authors needed to reclassify traits using expert knowledge given that the individual databases employed different coding approaches (i.e. European and New Zealand databases fuzzy coding, North American database binary coding). A centralized database with standardized and unambiguous traits and a consistent coding of traits would minimize the data processing effort. Discrepancies in the taxonomic resolutions (e.g. species, genus or family) when linking observational taxonomic data to trait database represents another challenge. When observations are on a more precise taxonomic level than data available in the trait databases (e.g. observations on species-level, trait data on genus-level) trait information of the less precise taxonomic level is often assigned (e.g. \cite{szocs_effects_2014, vos_taxonomic_2017}). Conversely, if trait information is only available on more precise taxonomic levels than the observed taxa, traits are aggregated to a less precise taxonomic level, e.g. \cite{poff_functional_2006, szocs_effects_2014, piliere_a._f._h._importance_2016, aspin_extreme_2019}. Studies have used different methods of trait aggregation, e.g. the mean \cite{magliozzi_functional_2019}, median \cite{szocs_effects_2014} or the mode \cite{piliere_a._f._h._importance_2016}. Up to now, studies on how and to which extent different trait aggregation methods influence trait-based analysis are missing. However, related knowledge would inform future studies regarding the choice of the aggregation method.

@@

@@

We used fuzzy coding for establishing our harmonized datasets unless data quality prohibited. In the latter case we used binary traits, i.e. categorical and continuous traits were converted into binary traits. For example, in the  freshwaterecology.info database, the classification of the trait voltinism accounts for different faunistic regions. Hence, entries such as "arctic" or "boreal" in the voltinism traits were substituted with a value of 1. Implicitly, we assumed for binary traits that a value of 1 and 0 corresponds to the highest and no affinity for a particular trait. Fuzzy codes are reported with different ranges in the trait databases (e.g. freshwaterecology.info 0 to 10, Tachet database 0 to 3 or 0 to 5). We standardized these to a range between 0 and 1 and converted trait affinities to percentages. Thus, fuzzy coded and binary traits were in the same range.

@@

To examine the influence of the taxonomic hierarchy and the trait variability on the outcome of the different trait aggregation methods we created three hypothetical examples of different taxonomic hierarchies (scenarios). 
1) A family with an equal number of genera and species (5 genera each with 5 species respectively), denoted as \textit{sim\_base}.
2) A family where one genus has a much larger number of species than the other 4 genera (1 genus with 13 species, 4 genera with 3 species respectively), denoted as \textit{sim\_extreme}. 
3) A family where all genera have a different number of species (8, 2, 7, 3, 5), denoted as \textit{sim\_variation}. In total every family consisted of 25 species. To each hypothetical taxonomic hierarchy a hypothetical grouping feature with 3 traits was assigned. The 25 affinities for each trait were simulated by sampling from a truncated normal distribution with a mean value of 0.5 and 5 levels of standard deviations (0.2, 0.4, 0.6, 0.8, and 1) respectively to simulate different levels of trait variability. The truncated normal distribution was bound to 0 and 1. Simulated trait affinities were converted to percentages, similar to the data processing of the trait databases. The sampling was repeated 100 times for each standard deviation. Hence, in total 100 datasets for each level of trait variability were simulated, resulting in 12500 simulated trait affinities for each scenario ($25$ species per family $* 5$ levels of trait variability $* 100$ replicates). All trait aggregation methods were applied to each simulated dataset resulting in 500 aggregated trait values per scenario. The results were compared based on the produced range of aggregated trait affinities between levels of trait variability and hypothetical taxonomic hierarchies. We also compared for each simulated dataset the differences in trait affinities obtained by each aggregation method.

@@

Definitions of grouping features and traits varied in their level of detail in the investigated trait databases. The freshwaterecology.info database, the Tachet database and the North American database (Twardochleb et al.) provided more detailed descriptions of their trait information compared to the North American (Vieira et al.) and New Zealand database. An exception is the Australian trait database which is a collection of seven specific trait datasets \cite{kefford_integrated_2020}. Thus, grouping features occur multiple times with varying differentiation into traits. Depending on the dataset trait information is described with more or less detail.

@@

Various codings of traits are used throughout the databases (e.g. binary, fuzzy, continuous). The freshwaterecology.info and Australian use different codings in their databases. Tachet and the New Zealand database exclusively use fuzzy coding. Both North American trait databases contain categorical grouping features that can be converted into traits using a binary coding (Table \ref{tab:trait_databases_coding_differentiation}).

@@

Comparing the aggregated trait affinities, in total 15000 ($500$ aggregated traits $*10$ unique comparisons of trait affinities and $3$ scenarios), for each simulated dataset showed that in most simulated datasets the differences between the aggregated traits were small. Only 1.42 \% (213 out of 15.000) of all comparisons showed a difference equal or greater than an absolute trait affinity of 0.1. Most of these differences occurred for the \textit{sim\_extreme} scenario (83.5 \%). The majority of the differences equal or above 0.1 were found between the aggregation methods \textit{direct\_agg\textsubscript{mean}} and \textit{stepwise\_agg\textsubscript{median}}, \textit{direct\_agg\textsubscript{median}} and \textit{stepwise\_agg\textsubscript{median}}, and \textit{stepwise\_agg\textsubscript{median}} and \textit{weighted\_agg} (Figure S\ref{fig:sim_indv_runs}). 
% Differences were rather small, largest difference 0.265 for sim_extreme (stepwise_median vs weighted_agg)

@@

\begin{figure}[H]
    \label{fig:rda_species_scores}
    \centering
    \includegraphics[width=16.5cm, height=10cm]{Species_scores_rda.png}
    \caption{Species scores obtained by RDA from the original analysis \cite{szocs_effects_2014}, using harmonized grouping features, and using harmonized grouping features with traits aggregated to family-level. The dots represent the individual species scores for each analysed trait along the conductivity axis. The violin plot shows the density estimate of the species scores. Gray vertical lines indicate the median of the obtained species scores.}
\end{figure}

@@

% Discussion/Comparison aggr. methods  
Our results show that aggregation of traits by the median compared to aggregating by the mean is often closer to the assigned traits at family level. However, in those cases where differences in the trait aggregation methods that use the median occurred, these differences were greater compared to the aggregation methods that used the mean. The different weighting methods, i.e. aggregating stepwise, direct or weighted, did not seem to have a strong impact, because the distribution of absolute trait affinity differences to assigned traits obtained by the \textit{direct\_agg\textsubscript{mean}}, \textit{stepwise\_agg\textsubscript{mean}} and \textit{weighted\_agg} were relatively similar in both tested datasets. This also applies to the distributions of absolute trait affinity differences obtained by \textit{direct\_agg\textsubscript{median}} and \textit{stepwise\_agg\textsubscript{median}}. Instead, using the mean or the median changed the distribution of differences in trait affinities between aggregated and assigned traits. A possible reason why the use of different weightings for taxa had little impact on trait aggregation is that a considerable portion of taxa used have a low numbers of genera or species entries. From the taxa that were compared from the North American trait dataset to the assigned traits at family-level, 13.8 \% were resolved on family-level and 61.7 \% were resolved on genus-level. 52.1 \% had 5 or less genera and 12.8 \% contained just one genus (Figure S\ref{fig:tax_hierarchy_NOA}). In the Australian dataset 21.4 \% of the compared taxa had trait information resolved on family-level, 39.5 \% resolved on genus level, and 39.1 \% resolved on species-level. 67.73 \% of the taxa contain 5 or less genera and 39.6 \% just contained one genus (Figure S\ref{fig:tax_hierarchy_AUS}). The simulation results indicated that the choice of the aggregation method for small numbers of genera per family (5) can be important when one genus has a much larger number of species compared to other genera within a particular family. % How often in the trait datasets?
In general however, inspecting the simulated datasets revealed that often aggregated traits obtained from the different aggregation methods only differ slightly.

% Differences between Results of AUS and NOA dataset
Observed differences between aggregated and assigned traits were higher for the North American trait dataset. This can be explained by the binary coding of the traits in the North American dataset. As a results, traits have been assigned affinities of either 1 or 0.  Therefore, traits of a certain grouping feature have a higher difference in trait affinities to each other than if they were coded with fuzzy codes.

% Discussion on differences between aggregated and assigned traits  
Evaluation of the differences between aggregated and assigned traits is difficult because it remains unclear what the true value of a particular trait for a particular family is. It is indeed questionable whether the aggregation of trait information on species or genus-level to a point estimate on family-level can be an adequate description of a particular trait at all, especially for traits with a high variability. In fact, traits can vary strongly on more precise taxonomic levels than the family-level. For example, Monaghan and Soares found a high dispersion of trait affinities for genera within families in the Tachet database for traits of the grouping features body size, flow preference, and reproduction \cite{monaghan_improving_2013}. Furthermore, studies focusing on the lability of traits, i.e. how much traits are unconstrained by phylogeny, found that traits of ecological preferences (e.g. thermal preference), body size, resistance forms, and to a lesser extent feeding mode are labile, and thus possibly highly variable \cite{poff_functional_2006,wilkes_traitbased_2020}. It is likely that assigned traits and aggregated traits at family-level will especially differ in affinities for the aforementioned grouping features. Our analysis could confirm this partially. The highest possible difference of 1 between aggregated trait affinities and assigned trait affinities (in total in 6.9 \% of all cases) for the Australian dataset occurred most often for body size large and medium (42.5 \% and 39.2 \% of all cases with a difference of 1, respectively) across all comparisons. Both traits had also the highest mean and standard deviation in the absolute difference of trait affinities across all tested comparisons (size medium: mean 0.5, sd 0.29; size large: mean 0.59, sd 0.32). For the taxa compared from the North American dataset, the highest possible difference in trait affinities occurred most often for the trait body size medium (15.8 \% of all cases with the maximum difference of 1), followed by the traits respiration via plastron and spiracle and respiration via tegument (both 13 \% of all cases with a maximum difference of 1). Respiration via plastron and spiracle showed the highest mean and standard deviation of the absolute difference of trait affinities across all comparisons (mean 0.71, sd 0.33), followed by respiration via tegument (mean 0.5, sd 0.34) and body size medium (mean 0.49, sd 0.3). 
% Check for those cases that deviate a lot -> how have the aggregation and assignments performed when looking back at the original data?

% Also briefly talk about traits that are more consistent?

% Include?:
% What do we really know because someone has watched the animals and published their behaviour; what is extrapolated knowledge? Where does the classification come from?
% Are the differences between continents because of evolution or only because scientists have done more research in one of the continents?

% Outlook: Are there alternatives to trait aggregation? Some statistical models combine trait information for priors (species distribution models/population growth models, see Kindvater et al 2018 and Brown & Roff)

% Further ideas: 
%   - Still some similarities between compared datasets -> Where does the classification come from?
%     Few sources used in NOA also used in AUS:
%     AUS: Merrit & Cummins 1978/1984 for feeding mode for a few genera and families 
%     NOA: Merritt, Richard W., Cummins, Berg 2008 -> but not clear if for feeding mode as well; according to trait description "Cummins, K. W. Trophic Relations of Aquatic Insects" was used for feeding mode
%   - Could not include all data due to data gaps -> much needed to be filled -> would maybe change the picture?

@@

The re-analysis showed that neither using harmonised grouping features nor aggregated traits changed drastically the results of the original study. The traits that did not respond to the salinity gradient (i.e. not on the extremes of the conductivity axis) in the re-analysis were those closest to the criterion that defined when a trait was associated with either high or low salinity (mahalanobis distance greater than the 97.5 \% quantile of the Chi-square distribution, see methods), namely feeding mode shredder, respiration gills, and life cycle duration traits. During harmonisation the trait shredder has been amalgamated based on three traits (Miner, xylophagus, and shredder). As a result, the trait affinities in the original dataset have a higher mean and standard deviation than in the aggregated and harmonized datasets (Table S\ref{tab:SI_resp_traits_summary_stats}). Hence, the signal in the original data for the trait shredder has presumably been weakened by the harmonization, which might be the reason that shredders showed a less strong reaction to the salinity gradient.

@@

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Supporting information %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{table}{0}
\setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}

### 2020-11-23 Paper/Results/SI.tex
23.11: Polishing/minor changes

\begin{table}[ht]
    \centering
    \caption{Mean, median and standard deviation of the traits that were responsive to the salinity gradient in the original study but not when re-analysis with the harmonized dataset trait dataset.} 
    \label{tab:SI_resp_traits_summary_stats}
    \begin{tabular}{llccc}
    \toprule[.1em]
    Dataset & Trait & Mean & Median & SD \\ 
    \toprule[.1em]
    Stepw\_median & Shredder & 0.20 & 0.14 & 0.25 \\ 
      Stepw\_mean & Shredder & 0.18 & 0.12 & 0.22 \\ 
      Direct\_median & Shredder & 0.21 & 0.14 & 0.25 \\ 
      Direct\_mean & Shredder & 0.19 & 0.14 & 0.22 \\ 
      Weighted & Shredder & 0.19 & 0.14 & 0.22 \\ 
      Harmonized; not\_aggregated & Shredder & 0.18 & 0.12 & 0.24 \\ 
      Original & Shredder & 0.25 & 0.14 & 0.32 \\ 
      Stepw\_median & Gills & 0.30 & 0.27 & 0.32 \\ 
      Stepw\_mean & Gills & 0.29 & 0.22 & 0.32 \\ 
      Direct\_median & Gills & 0.30 & 0.30 & 0.32 \\ 
      Direct\_mean & Gills & 0.30 & 0.30 & 0.32 \\ 
      Weighted & Gills & 0.30 & 0.30 & 0.32 \\ 
      Harmonized; not\_aggregated & Gills & 0.30 & 0.25 & 0.32 \\ 
      Original & Gills & 0.28 & 0.00 & 0.33 \\ 
      Stepw\_median & Short life cycle & 0.64 & 0.75 & 0.39 \\ 
      Stepw\_mean & Short life cycle & 0.64 & 0.79 & 0.39 \\ 
      Direct\_median & Short life cycle & 0.67 & 0.75 & 0.37 \\ 
      Direct\_mean & Short life cycle & 0.67 & 0.79 & 0.38 \\ 
      Weighted & Short life cycle & 0.67 & 0.79 & 0.38 \\ 
      Harmonized; not\_aggregated & Short life cycle & 0.64 & 0.75 & 0.40 \\ 
      Original & Short life cycle & 0.64 & 0.75 & 0.40 \\ 
      Stepw\_median & Long life cylce & 0.36 & 0.25 & 0.39 \\ 
      Stepw\_mean & Long life cylce & 0.36 & 0.21 & 0.39 \\ 
      Direct\_median & Long life cylce & 0.33 & 0.25 & 0.37 \\ 
      Direct\_mean & Long life cylce & 0.33 & 0.21 & 0.38 \\ 
      Weighted & Long life cylce & 0.33 & 0.21 & 0.38 \\ 
      Harmonized; not\_aggregated & Long life cylce & 0.36 & 0.25 & 0.40 \\ 
      Original & Long life cylce & 0.36 & 0.25 & 0.40 \\ 
    \bottomrule
    \end{tabular}
\end{table}

>>>>

\begin{table}[ht]
    \centering
    \caption{Mean, median and standard deviation of affinities of the traits that were responsive to the salinity gradient in the original study but not when re-analysis with the harmonized dataset trait dataset.} 
    \label{tab:SI_resp_traits_summary_stats}
    \begin{tabular}{llccc}
    \toprule[.1em]
    Dataset & Trait & Mean & Median & SD \\ 
    \toprule[.1em]
    Stepw\_median & Shredder & 0.20 & 0.14 & 0.25 \\ 
      Stepw\_mean & Shredder & 0.18 & 0.12 & 0.22 \\ 
      Direct\_median & Shredder & 0.21 & 0.14 & 0.25 \\ 
      Direct\_mean & Shredder & 0.19 & 0.14 & 0.22 \\ 
      Weighted & Shredder & 0.19 & 0.14 & 0.22 \\ 
      Harmonized; not\_aggregated & Shredder & 0.18 & 0.12 & 0.24 \\ 
      Original & Shredder & 0.25 & 0.14 & 0.32 \\ 
      Stepw\_median & Gills & 0.30 & 0.27 & 0.32 \\ 
      Stepw\_mean & Gills & 0.29 & 0.22 & 0.32 \\ 
      Direct\_median & Gills & 0.30 & 0.30 & 0.32 \\ 
      Direct\_mean & Gills & 0.30 & 0.30 & 0.32 \\ 
      Weighted & Gills & 0.30 & 0.30 & 0.32 \\ 
      Harmonized; not\_aggregated & Gills & 0.30 & 0.25 & 0.32 \\ 
      Original & Gills & 0.28 & 0.00 & 0.33 \\ 
      Stepw\_median & Short life cycle & 0.64 & 0.75 & 0.39 \\ 
      Stepw\_mean & Short life cycle & 0.64 & 0.79 & 0.39 \\ 
      Direct\_median & Short life cycle & 0.67 & 0.75 & 0.37 \\ 
      Direct\_mean & Short life cycle & 0.67 & 0.79 & 0.38 \\ 
      Weighted & Short life cycle & 0.67 & 0.79 & 0.38 \\ 
      Harmonized; not\_aggregated & Short life cycle & 0.64 & 0.75 & 0.40 \\ 
      Original & Short life cycle & 0.64 & 0.75 & 0.40 \\ 
      Stepw\_median & Long life cylce & 0.36 & 0.25 & 0.39 \\ 
      Stepw\_mean & Long life cylce & 0.36 & 0.21 & 0.39 \\ 
      Direct\_median & Long life cylce & 0.33 & 0.25 & 0.37 \\ 
      Direct\_mean & Long life cylce & 0.33 & 0.21 & 0.38 \\ 
      Weighted & Long life cylce & 0.33 & 0.21 & 0.38 \\ 
      Harmonized; not\_aggregated & Long life cylce & 0.36 & 0.25 & 0.40 \\ 
      Original & Long life cylce & 0.36 & 0.25 & 0.40 \\ 
    \bottomrule
    \end{tabular}
\end{table}
